Corotating Clumps Around Adolescent Low-Mass Stars: Four Years of TESS

### 2023-08-06 paper/ms.tex
incorporate Luisa Rebull's suggestions

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% MS TODO's
%
% * find/replace "centrifugally supported" with something less
% jargony, and more physically meaningful.  (read CR89)
%
% * "Bound-free absorption can provide
%   wavelength-dependent opacity, but the absorption coefficients
%   generally {\it grow} with increasing wavelength, rather than {\it
%   shrink} as one needs in order to get deeper absorption in the blue
%   than in the red."
%   --> Isn't this statement wavelength dependent????
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

All pre-main-sequence stars vary in optical brightness, and the origin
of such variability is, in most cases, understood.  Well-explored
sources of optical variability include inhomogeneities on stellar
surfaces such as starspots and faculae \citep{2021isma.book.....B},
occultations by gas-rich circumstellar disks
\citep{2017MNRAS.470..202B}, and, in geometrically favorable
circumstances, eclipses by stars and planets
\citep{2010exop.book...55W}.  More exotic forms of optical variability
relevant to this work include transiting exocomets (e.g.~$\beta$~Pic;
\citealt{2019A&A...625L..13Z}) and disintegrating rocky bodies around
both M dwarfs (e.g. KOI-2700; \citealt{2014ApJ...784...40R}) and white
dwarfs (e.g. WD~1145; \citealt{2015Natur.526..546V}).

@@

Once a star's period $P$ was identified, we binned the phased light
curve to 100 points per cycle.  To separate ``sharp'' local minima
from smooth spot-induced variability, we then iteratively fit
penalized splines to the wrapped phase-folded light curve, excluding
points more than two standard deviations away from the local continuum
\citep{2019AJ....158..143H}.  The maximum number of equidistant spline
knots per cycle is the parameter in this framework that controlled the
meaning of ``sharp'' --- we allowed at most 10 such knots per cycle,
though for most stars fewer knots were preferred based on an
$\ell^2$-norm penalty.

@@

For the SED fitting, we used \texttt{astroARIADNE}
\citep{2022MNRAS.513.2719V}.  We adopted the BT-Settl stellar
atmosphere models \citep{Allard2012} assuming the
\citet{2009ARA&A..47..481A} solar abundances, and the
\citet{2006MNRAS.368.1087B} water line lists.  The broadband
magnitudes we considered included $GG_{\rm BP}G_{\rm RP}$ from Gaia
DR2, $Vgri$ from APASS, $JHK_{\rm S}$ from 2MASS, SDSS $riz$, and the
WISE 1-2 passbands.  We specifically omitted UV flux measurements to
avoid biasing our fit with any possible chromospheric UV excess.
\texttt{astroARIADNE} compares the measured broadband flux
measurements against pre-computed model grids, and by default fits for
six parameters: $\{ T_{\rm eff}, R_\star, A_{\rm V}, \log g, [{\rm
Fe/H}], d \}$.  The distance  prior is drawn from
\citet{2021AJ....161..147B}.  The surface gravity and metallicity are
generally unconstrained.  And finally, given our particular use-case,
we assumed the following priors for the temperature, stellar size, and
extinction:
\begin{align}
  T_{\rm eff} / {\rm K}    &\sim \mathcal{N}(3000, 1000), \\
  R_\star / R_\odot  &\sim \mathcal{T}_{\rm N}(0.5, 0.3, 0.1, 1.5), \\
  A_{\rm V} / {\rm mag}    &\sim \mathcal{U}(0, 0.2),
\end{align}
for $\mathcal{N}$ the Gaussian and $\mathcal{U}$ the uniform
distributions, and $\mathcal{T}(\mu, \sigma, a, b)$ a truncated normal
distribution with mean $\mu$, standard deviation $\sigma$, and lower
and upper bounds $a$ and $b$.  Using \texttt{Dynesty}
\citep{2020MNRAS.493.3132S}, we statically sampled the posterior
probability assuming the default Gaussian likelihood, and set a
stopping threshold of ${\rm d}\log \mathcal{Z} < 0.01$, where
$\mathcal{Z}$ denotes the evidence.

@@

\begin{figure*}[!tp]
	\begin{center}
		\centering
		\includegraphics[width=0.98\textwidth]{f3.pdf}
    \vspace{-0.3cm}
		\caption{
      {\bf CQVs found in the TESS 2-minute data.}
      Phased TESS light curves over one month are shown for \ngoods\
      CQVs.  Gray are raw 2-minute data; black bins to 300 points per
      cycle.  Each panel is labeled by the TIC identifier, the TESS
      sector number, and the period in hours.  Objects are ordered
      such that sources with the most TESS data available are on top
      (see Section~\ref{sec:catalog}).
		}
		\label{fig:cqvs}
	\end{center}
\end{figure*}

@@

Table~\ref{tab:thetable} lists the \ncpvsfound\ objects identified by
our search.  \ngoods\ objects demonstrated what we viewed as
unambiguous characteristics of the CQV phenomenon in at least one TESS
sector; we refer to these as the ``high-quality sample''.  The
classification of the remaining \nmaybes\ CQV candidates was
ambiguous.  Additional data from TESS or other instruments could help
resolve their classification.  The boolean \texttt{quality} column in
the table divides the two classes.  In the following, we will restrict
our discussion to the high-quality sample.

@@

The two ``good'' IR excesses belonged to ``ambiguous'' CQVs
TIC~193136669 (TWA~34) and TIC~57830249 (TWA~33).  Both are in the TW
Hydrae association ($\approx$10\,Myr), and have relatively long
periods of 38\,hr and 44\,hr respectively.  In our initial labeling,
we labeled both as ``ambiguous'' CQVs because the dips in their light
curves did not show the rigid periodicity typical of CQVs;  their
periods were also relatively long.  After labeling them, inspection
of further sectors clarified that both sources are dippers (see plots
in Figure~\ref{fig:vet}).  Independently, TIC 193136669 has a cold
molecular disk based on observed 1.3\,mm continuum emission and
resolved Keplerian $^{12}{\rm CO}(2-1)$ emission
\citep{2015A&A...582L...5R}.  It was also labeled a dipper by
\citet{2022ApJS..263...14C}; we agree with their designation, but
nonetheless leave it in our catalog as an indication of possible
ambiguities in our search process.  TIC 57830249 (TWA~33) also has
previously detected 1.3\,mm continuum emission
\citep{2015A&A...582L...5R}, suggestive of cold dust grains being
present.  It is also a dipper.
Section~\ref{subsec:discdippers} highlights plausible evolutionary
connections between CQVs and dippers in light of these
``misclassifications'' .

Our three ``possible'' infrared excesses were TIC~405910546,
289840926, and 244161191.  After a literature search, we concluded
that none have clear evidence for the presence of a disk.
TIC~405910546, in LCC, shows a unique TESS light curve, reminiscent of
a $P$=38\,hr singly-eclipsing binary, except with additional
substructure during each eclipse that resembles the CQV phenomenon
more than any other variability of which we are aware.  TIC~289840926
($\beta$~Pic moving group, $P$=4.8\,hr), shows what we believe is a
clear CQV signal, but has no definitive evidence for a large, dusty
disk.  TIC~244161191 (TOI-278), in Columba, also has no
definitive evidence for a large disk.  It is however
``multi-periodic''---in addition to the 7.17\,hr CQV signal, this
source shows a superposed 8.39\,hr sinusoidal signal, probably from an
unresolved neighboring star.

@@

\begin{figure*}[!tp]
	\begin{center}
		\centering
		\includegraphics[width=\textwidth]{f4.pdf}
		\vspace{-0.6cm}
		\caption{
		{\bf CQVs keep their periods but change their shapes.}
      32 of our \ngoods\ CQVs from Figure~\ref{fig:cqvs} had
      120-second cadence TESS data available for a baseline of at
      least two years; the 27 brightest are shown here.  Each panel
      shows one sector of TESS data, and is phased to its deepest
      minimum in flux.  Each panel's title shows the TIC identifier
      and approximate period in hours.  Text insets show the TESS
      sector numbers, which generally span two years, or at least
      1{,}000 cycles.  The vertical scale is fixed across sectors to
      clarify shape changes.
      Gray circles are raw 2-minute data; colored circles bin to 300
      points per cycle. 
		}
		\label{fig:evoln}
	\end{center}
\end{figure*}

@@

\begin{figure*}[!t]
	\begin{center}
		\centering
		\includegraphics[width=\textwidth]{f6.pdf}
		\vspace{-0.6cm}
		\caption{
      {\bf Evolution of LP 12-502} ($P$=18.6\,h) at fixed period and
      epoch over three years.  Each panel shows one (stacked) TESS
      orbit; small text denotes relative cycle number.  There are 200
      binned black points per cycle.  The TESS pointing law dictates
      time gaps; larger gaps tend to yield larger shape changes.  The
      dips usually evolve over tens to hundreds of cycles.  However
      cycles 1233-1264 show a dip that ``switched'' from a depth and
      duration of 3\% and 3\,hr to 0.3\% and 1\,hr over less than one
      cycle (cf.~Figure~\ref{fig:lplc}).
		}
		\label{fig:lp}
	\end{center}
\end{figure*}

\subsubsection{LP 12-502 observations}
While many CQVs had multiple sectors of adjacent or nearly-adjacent
data, LP 12-502 (TIC 402980664; $d$=21\,pc, $J$=9.4, $T$=11.1) stood
out due to the quality and content of its data.  Figure~\ref{fig:lplc}
shows all available data from Sectors 18, 19, 25, 26, 53, 58, and 59,
split into successive orbits; the star was observed at 120-second
cadence whenever it was observable by TESS.  We binned the light curve
to 15-minute intervals for visual clarity, and required at least one
(120-second cadence) flux measurement per bin.  Points more than
2.5$\sigma$ above the median are drawn in gray, also for visual
clarity.  Missing data are not drawn.  Figure~\ref{fig:lp} then shows
the same data, but stacked into successive TESS orbits spanning half a
lunar month each.

@@

The second dip complex during cycles 248-315 shows the most
substructure.  During e.g.~cycles 283-298, this single complex shows
six local minima.  The deepest dip is sharp: it shows a flux excursion
of 3.5\% over about 22 minutes (0.02\,$P$), which is the steepest
slope exhibited anywhere in the LP~12-502 remarkable dataset.  After
the sharp dip, there is a roughly exponential fall-off spanning about
a quarter of a period, punctuated by coherent local minima and maxima
which in detail (Appendix~\ref{app:lp}) have slightly longer periods
than the sharp dip.  The sharp leading dip only decreases in depth
following a sudden state-switch at BTJD 2030.7 (cycles 299-315), which
happens during a flare (Figure~\ref{fig:lplc}).  The trailing dips
remain thereafter.

@@

\subsubsection{Lessons from LP 12-502}

@@

\begin{figure*}[!t]
	\begin{center}
		\centering
		\includegraphics[width=0.9\textwidth]{f7.pdf}
		\vspace{-0.1cm}
		\caption{
      {\bf Properties of CQVs identified by our search}.
      The \ngoods\ bonafide CQVs in Table~\ref{tab:thetable} are the
      dark blue circles; \nmaybes\ ambiguous CQV candidates are light
      blue circles.  The top panels show the \nstarssearched\ target
      stars with 120-second cadence TESS data as the shaded gray
      background; darker regions correspond to a larger relative
      number of searched stars.  The lower-left panel compares the
      rotation--color distribution of CQVs against the rotation
      periods of K and M dwarfs in the Pleiades from
      \citet{2016AJ....152..114R}.  The lower-middle panel plots the
      derived corotation radii $R_{\rm cr} = (GM/\Omega^2)^{1/3}$ in
      units of stellar radii against the measured CQV periods, in
      units of hours.
		}
		\label{fig:catalogscatter}
	\end{center}
\end{figure*}

@@

The most massive CQVs in our sample might represent an interesting new
regime.  To date, the only stars reported to show the CQV phenomenon
are M dwarfs, with typical stellar masses $\lesssim$0.3\,$M_\odot$
\citep{2022AJ....163..144G}.  TIC~405754448 and TIC~405910546 however
appear to have masses of $\approx$0.82 and $\approx$0.60\,$M_\odot$
respectively.  The next-highest masses are $\approx$0.40\,$M_\odot$.
The masses for the former two objects are consistent with their CAMD
locations, and their membership in LCC.  Based on its light curve
morphology, TIC~405910546 should be studied in greater depth, to
confirm it is not an eclipsing binary.  TIC~405754448 similarly shows
distinct morphology from many of the CQVs in Figure~\ref{fig:cqvs}, in
that it has some of the lowest amplitude dips.  While TIC~405754448
might be an unresolved binary (RUWE$_{\rm DR3}$=4.7),  the depth of
the coherent dips present in the Sector~11 light curve imply that the
CQV signal is indeed from the apparent primary of the system.  Taken
in sum, both of these objects seem to suggest that the CQV phenomenon
extends up in mass to pre-main-sequence K dwarfs.

@@

\subsubsection{The lifetime constraint}
Independent of the microphysical basis for gas absorption, the
observed lifetime of the CQV phenomenon could provide another
dimension to discern between the gas {\it vs.} dust clump scenarios.
Based on the available statistics from the studies by L.~Rebull and
J.~Stauffer, it seems plausible that CQV occurrence decreases in time,
bottoming to zero well before the $\approx$700\,Myr age of Praesepe.
This is odd in the context of the prominence scenario, because
pre-main-sequence M dwarfs spin up over the first $\approx$$10^8$\,yr;
prominences might even be expected to be {\it more} common at the age
of the Pleiades than for younger stars.  This broadly assumes that any
star that can support prominences will, which translates in detail to
a condition on the star's mass, rotation period, and magnetic field
strength.  Based on the degree to which M dwarfs remain rapid rotators
over the first 700\,Myr \citep[e.g. Figure~13
of][]{2022AJ....164...80R}, it seems surprising in the gas clump
(prominence) scenario that very few stars older than 150\,Myr have
been seen to exhibit the phenomenon.  In the dust clump scenario this
is not a problem, because if the dust were externally sourced, it
would be expected to have a finite supply.

@@

The interval of half a cycle period could be significant in the
context of CQVs for two reasons.  The first is that for material on a
circular orbit viewed edge-on, it corresponds to the interval between
transit and secondary eclipse.  The second is that it also corresponds
to the interval over which half of the star's surface is visible.  Of
the CQVs in Figures~\ref{fig:cqvs} and~\ref{fig:evoln}, a number that
seems greater than random might exhibit a preference for showing dips
or peaks that correspond to the half-cycle interval.

@@

In this work, we searched 120-second cadence TESS data collected
between July~2018 and Sep~2022 for complex quasiperiodic variables
(CQVs).  Our search sample included \nstarssearched\ K and M dwarfs
within 150\,pc, and was $\gtrsim 80\%$ complete within 30\,pc, and
$\lesssim 10\%$ complete at distances exceeding 100\,pc.

@@

The distinction between gas or dust is important because it will
clarify whether the CQV phenomenon is intrinsic, so that the material
comes from the star, or extrinsic, so that it is sourced through some
generic evolutionary phase of circumstellar disks.  This knowledge would in
turn propagate to our understanding of whether observing the
phenomenon is primarily teaching us about e.g.~dust production and
processing in circumstellar disks, or whether it is teaching us about
the ability of cold gas to remain stable in hot stellar coronae for
long durations.
Observationally, acquisition of medium- or high-resolution time-series
spectroscopy holds the best chance at resolving the gas vs.~dust
question.  Given the observed evolutionary timescales, such data
should be acquired simultaneously with photometric time-series
observations to help clarify its interpretation.

@@

\begin{figure*}[!t]
	\begin{center}
    \includegraphics[width=0.45\textwidth]{ANNOTATED_resid_TIC_402980664_P18.5611_2min_phase_timegroups.png}
    	\end{center}
    \vspace{-0.4cm}
		\caption{
	      {\bf Alternative view of the evolution of LP 12-502}
	      (Figure~\ref{fig:lp}), arranged to emphasize changes in transit
	      times.  There are 200 binned black points per cycle; a two-harmonic
	      sinusoid has been subtracted over specific chunks in time ({\bf see text}).
	      Vertical gray lines are underplotted to help guide the eye to instances
	      in which preferred dip phases synchronize over long baselines.
	      The orange and green lines guide the eye to where dips
	      appear to change the positions of their local minima.
		}
		\label{fig:lp2}
\end{figure*}

@@

\section{More river plots}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% MS TODO's
%
% * find/replace "centrifugally supported" with something less
% jargony, and more physically meaningful.  (read CR89)
%
% * "Bound-free absorption can provide
%   wavelength-dependent opacity, but the absorption coefficients
%   generally {\it grow} with increasing wavelength, rather than {\it
%   shrink} as one needs in order to get deeper absorption in the blue
%   than in the red."
%   --> Isn't this statement wavelength dependent????
%
% * Verify "32" in Figure 4 does not change with number of stars in
%   subsample changing
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

All pre-main-sequence stars vary in optical brightness, and the origin
of such variability is, in most cases, understood.  Well-explored
sources of optical variability include inhomogeneities on stellar
surfaces such as starspots and faculae \citep[e.g.][]{2021isma.book.....B},
occultations by gas-rich circumstellar disks
\citep[e.g.][]{2017MNRAS.470..202B}, and, in geometrically favorable
circumstances, eclipses by stars and planets
\citep[e.g.][]{2010exop.book...55W}.  More exotic forms of optical variability
relevant to this work include transiting exocomets (e.g.~$\beta$~Pic;
\citealt{2019A&A...625L..13Z}) and disintegrating rocky bodies around
both M dwarfs (e.g. KOI-2700; \citealt{2014ApJ...784...40R}) and white
dwarfs (e.g. WD~1145; \citealt{2015Natur.526..546V}).

@@

Once a star's period $P$ was identified, we binned the phased light
curve to 100 points per cycle.  To separate ``sharp'' local minima
from smooth spot-induced variability, we then iteratively fit robust
penalized splines to the wrapped phase-folded light curve, excluding
points more than two standard deviations away from the local continuum
\citep{2019AJ....158..143H}.  The maximum number of equidistant spline
knots per cycle is the parameter in this framework that controlled the
meaning of ``sharp'' --- we allowed at most 10 such knots per cycle,
though for most stars fewer knots were preferred based on
cross-validation using an $\ell^2$-norm penalty.  An example fit is
shown in panel {\it (e)} of Figure~\ref{fig:vetting}.

@@

For the SED fitting, we used \texttt{astroARIADNE}
\citep{2022MNRAS.513.2719V}.  We adopted the BT-Settl stellar
atmosphere models \citep{Allard2012} assuming the
\citet{2009ARA&A..47..481A} solar abundances, and the
\citet{2006MNRAS.368.1087B} water line lists.  The broadband
magnitudes we considered included $GG_{\rm BP}G_{\rm RP}$ from Gaia
DR2, $Vgri$ from APASS, $JHK_{\rm S}$ from 2MASS, SDSS $riz$, and the
WISE 1-2 passbands.  We specifically omitted UV flux measurements from
our SED fit to avoid any possible bias induced by chromospheric UV
excess; we similarly omitted WISE 3-4, but due to reliability
concerns.
\texttt{astroARIADNE} compares the measured broadband flux
measurements against pre-computed model grids, and by default fits for
six parameters: $\{ T_{\rm eff}, R_\star, A_{\rm V}, \log g, [{\rm
Fe/H}], d \}$.  The distance  prior is drawn from
\citet{2021AJ....161..147B}.  The surface gravity and metallicity are
generally unconstrained.  And finally, given our particular use-case,
we assumed the following priors for the temperature, stellar size, and
extinction:
\begin{align}
  T_{\rm eff} / {\rm K}    &\sim \mathcal{N}(3000, 1000), \\
  R_\star / R_\odot  &\sim \mathcal{T}_{\rm N}(0.5, 0.3, 0.1, 1.5), \\
  A_{\rm V} / {\rm mag}    &\sim \mathcal{U}(0, 0.2),
\end{align}
for $\mathcal{N}$ the Gaussian and $\mathcal{U}$ the uniform
distributions, and $\mathcal{T}(\mu, \sigma, a, b)$ a truncated normal
distribution with mean $\mu$, standard deviation $\sigma$, and lower
and upper bounds $a$ and $b$.  Using \texttt{Dynesty}
\citep{2020MNRAS.493.3132S}, we statically sampled the posterior
probability assuming the default Gaussian likelihood, and set a
stopping threshold of ${\rm d}\log \mathcal{Z} < 0.01$, where
$\mathcal{Z}$ denotes the evidence.

@@

\begin{figure*}[!tp]
	\begin{center}
		\centering
		\includegraphics[width=0.98\textwidth]{f3.pdf}
    \vspace{-0.3cm}
		\caption{
      {\bf CQVs found in the TESS 2-minute data.}
      Phased TESS light curves over one month are shown for \ngoods\
      CQVs in the high quality sample.  Gray are raw 2-minute data;
      black bins to 300 points per cycle.  Each panel is labeled by
      the TIC identifier, the TESS sector number, and the period in
      hours.  Objects are ordered such that sources with the most TESS
      data available are on top (see Section~\ref{sec:catalog}).
		}
		\label{fig:cqvs}
	\end{center}
\end{figure*}

@@

Table~\ref{tab:thetable} lists the \ncpvsfound\ objects identified by
our search.  The ``high-quality sample'' includes \ngoods\ of these
objects, which demonstrated what we viewed as
unambiguous characteristics of the CQV phenomenon in at least one TESS
sector.  The
classification of the remaining \nmaybes\ CQV candidates was
ambiguous.  Additional data from TESS or other instruments could help
resolve their classification.  The boolean \texttt{quality} column in
the table divides the two classes.  In the following, we will restrict
our discussion to the high-quality sample.

@@

The two ``good'' IR excesses belonged to ``ambiguous'' CQVs
TIC~193136669 (TWA~34) and TIC~57830249 (TWA~33).  Both are in the TW
Hydrae association ($\approx$10\,Myr), and have relatively long
periods of 38\,hr and 44\,hr respectively.  In our initial labeling,
we labeled both as ``ambiguous'' CQVs because the dips in their light
curves did not show the rigid periodicity typical of CQVs;  their
periods were also relatively long.  After labeling them, inspection
of further sectors clarified that both sources are dippers (see online plots
in Figure~\ref{fig:vet}).  Independently, TIC 193136669 has a cold
molecular disk based on observed 1.3\,mm continuum emission and
resolved Keplerian $^{12}{\rm CO}(2-1)$ emission
\citep{2015A&A...582L...5R}.  It was also labeled a dipper by
\citet{2022ApJS..263...14C}; we agree with their designation, but
nonetheless leave it in our catalog as an indication of possible
ambiguities in our search process.  TIC 57830249 (TWA~33) also has
previously detected 1.3\,mm continuum emission
\citep{2015A&A...582L...5R}, suggestive of cold dust grains being
present.  It is also a dipper.
Section~\ref{subsec:discdippers} highlights plausible evolutionary
connections between CQVs and dippers in light of these
``misclassifications'' .

Our three ``possible'' infrared excesses were TIC~405910546,
289840926, and 244161191.  After a literature search, we concluded
that none have clear evidence for the presence of a disk.
TIC~405910546, in LCC, shows a unique TESS light curve, reminiscent of
a $P$=38\,hr singly-eclipsing binary, except with additional
substructure during each eclipse that resembles the CQV phenomenon
more than any other variability of which we are aware.  TIC~289840926
($\beta$~Pic moving group, $P$=4.8\,hr), shows what we believe is a
clear CQV signal, but has no definitive evidence for a large, dusty
disk.  TIC~244161191 (TOI-278), in Columba, also has no
definitive evidence for a large disk.  It is however
``multi-periodic''---in addition to the 7.17\,hr CQV signal, this
source shows a superposed 8.39\,hr sinusoidal signal, probably from an
unresolved neighboring star or binary companion.

@@

\begin{figure*}[!tp]
	\begin{center}
		\centering
		\includegraphics[width=\textwidth]{f4.pdf}
		\vspace{-0.6cm}
		\caption{
      %TODO : 
		{\bf CQVs keep their periods but change their shapes.}
      Out of the \ngoods\ CQVs in Figure~\ref{fig:cqvs}, 32 had
      120-second cadence TESS data available for a baseline of at
      least two years; the 27 brightest are shown here due to space
      constraints.  Each panel shows one sector of TESS data, and is
      phased to its deepest minimum in flux.  Each panel's title shows
      the TIC identifier and approximate period in hours.  Text insets
      show the TESS sector numbers, which generally span two years, or
      at least 1{,}000 cycles.  The vertical scale is fixed across
      sectors to clarify shape changes.  Gray circles are raw 2-minute
      data; colored circles bin to 300 points per cycle. 
		}
		\label{fig:evoln}
	\end{center}
\end{figure*}

@@

\begin{figure*}[!t]
	\begin{center}
		\centering
		\includegraphics[width=\textwidth]{f6.pdf}
		\vspace{-0.6cm}
		\caption{
      {\bf Evolution of LP 12-502} ($P$=18.6\,h) at fixed period and
      epoch over three years.  Each panel shows one (stacked) TESS
      orbit; small text denotes relative cycle number.  There are 200
      binned black points per cycle.  The TESS pointing law dictates
      the large time gaps between Cycles 64-248, 315-1233, and
      1264-1410; larger gaps tend to yield larger shape changes.  The
      dips usually evolve over tens to hundreds of cycles.  However
      cycles 1233-1264 show a dip that ``switched'' from a depth and
      duration of 3\% and 3\,hr to 0.3\% and 1\,hr over less than one
      cycle (cf.~Figure~\ref{fig:lplc}).
		}
		\label{fig:lp}
	\end{center}
\end{figure*}

Out of the \ngoods\ CQVs, a small fraction were in regions of sky with
many sectors of adjacent TESS data.  Out of these objects, LP 12-502
(TIC 402980664; $d$=21\,pc, $J$=9.4, $T$=11.1) stood out due to the
quality and content of its data.  We describe another interesting
source, TIC~300651846, in Appendix~\ref{app:tic3006}.  For this
section, we first present the LP 12-502 observations
(Section~\ref{subsec:lpobservations}), and then highlight their
interpretation (Section~\ref{subsec:lplessons}).

\subsubsection{LP 12-502 observations}
\label{subsec:lpobservations}

Figure~\ref{fig:lplc} shows all available data for LP 12-502 from TESS
Sectors 18, 19, 25, 26, 53, 58, and 59, split into successive orbits.
The star was observed at 120-second cadence whenever it was observable
by TESS.  We binned the light curve to 15-minute intervals for visual
clarity, and required at least one (120-second cadence) flux
measurement per bin.  Points more than 2.5$\sigma$ above the median
are drawn in gray, also for visual clarity.  Missing data are not
drawn.  Figure~\ref{fig:lp} then shows the same data, but stacked into
successive TESS orbits spanning half a lunar month each.

@@

The second dip complex during cycles 248-315 shows the most
substructure.  During e.g.~cycles 283-298, this single complex shows
six local minima.  The deepest dip is sharp: it shows a flux excursion
of 3.5\% over about 22 minutes (0.02\,$P$), which is the steepest
slope exhibited anywhere in the LP~12-502 dataset.  After
the sharp dip, there is a roughly exponential fall-off spanning about
a quarter of a period, punctuated by coherent local minima and maxima
which in detail (Appendix~\ref{app:lp}) have slightly longer periods
than the sharp dip.  The sharp leading dip only decreases in depth
following a sudden state-switch at BTJD 2030.7 (cycles 299-315), which
happens during a flare (Figure~\ref{fig:lplc}).  The trailing dips
remain thereafter.

@@

\subsubsection{Lessons from LP 12-502}
\label{subsec:lplessons}

@@

\begin{figure*}[!t]
	\begin{center}
		\centering
		\includegraphics[width=0.9\textwidth]{f7.pdf}
		\vspace{-0.1cm}
		\caption{
      {\bf Properties of CQVs identified by our search}.
      CQVs are pre-main-sequence K and M dwarfs, younger than
      $\approx$200 Myr, with rotation periods faster than around half
      of Pleiades-age M dwarfs.
      The \ngoods\ bonafide CQVs in Table~\ref{tab:thetable} are the
      dark blue circles; \nmaybes\ ambiguous CQV candidates are light
      blue circles.  The top panels show the \nstarssearched\ target
      stars with 120-second cadence TESS data as the shaded gray
      background; darker regions correspond to a larger relative
      number of searched stars.  The lower-left panel compares the
      rotation--color distribution of CQVs against the rotation
      periods of K and M dwarfs in the Pleiades from
      \citet{2016AJ....152..114R}.  The lower-middle panel plots the
      derived corotation radii $R_{\rm cr} = (GM/\Omega^2)^{1/3}$ in
      units of stellar radii against the measured CQV periods, in
      units of hours.  Ages in the final panel are known from cluster
      membership.
		}
		\label{fig:catalogscatter}
	\end{center}
\end{figure*}

@@

The most massive CQVs in our sample might represent an interesting new
regime.  To date, the only stars reported to show the CQV phenomenon
are M dwarfs, with typical stellar masses $\lesssim$0.3\,$M_\odot$
\citep{2017AJ....153..152S,2022AJ....163..144G}.  TIC~405754448 and TIC~405910546 however
appear to have masses of $\approx$0.82 and $\approx$0.60\,$M_\odot$
respectively.  The next-highest masses are $\approx$0.40\,$M_\odot$.
The masses for the former two objects are consistent with their
locations in the color--absolute magnitude diagram, and with their
membership in LCC.  Based on its light curve morphology, TIC~405910546
should be studied in greater depth, to confirm it is not an eclipsing
binary.  TIC~405754448 similarly shows distinct morphology from many
of the CQVs in Figure~\ref{fig:cqvs}, in that it has some of the
lowest amplitude dips.  While TIC~405754448 might be an unresolved
binary (RUWE$_{\rm DR3}$=4.7),  the depth of the coherent dips present
in the Sector~11 light curve imply that the CQV signal is indeed from
the apparent primary of the system.  Taken in sum, both of these
objects seem to suggest that the CQV phenomenon extends up in mass to
pre-main-sequence K dwarfs.

@@

\subsubsection{The lifetime constraint}
Independent of the microphysical basis for gas absorption, the
observed lifetime of the CQV phenomenon could provide another
dimension to discern between the gas {\it vs.} dust clump scenarios.
Based on the available statistics from e.g.
\citet{2022AJ....164...80R} and references therein, it seems plausible
that CQV occurrence decreases in time, bottoming to zero well before
the $\approx$700\,Myr age of Praesepe.  This is odd in the context of
the prominence scenario, because
pre-main-sequence M dwarfs spin up over the first $\approx$$10^8$\,yr;
prominences might even be expected to be {\it more} common at the age
of the Pleiades than for younger stars.  This broadly assumes that any
star that can support prominences will, and that the magnetic topology
of rapidly rotating M dwarfs at 100\,Myr is similar to that of 
rapidly rotating M dwarfs at say, 10\,Myr.  Based on the degree to
which M dwarfs remain rapid rotators over the first 700\,Myr
\citep[e.g. Figure~13 of][]{2022AJ....164...80R}, it seems surprising
in the gas clump (prominence) scenario that very few stars older than
150\,Myr have been seen to exhibit the phenomenon.  In the dust clump
scenario this is not a problem, because if the dust were externally
sourced, it would be expected to have a finite supply.

@@

The interval of half a cycle period could be significant in the
context of CQVs for two reasons.  The first is that for material on a
circular orbit viewed edge-on, a half-cycle corresponds to the interval between
transit and secondary eclipse.  The second is that the half-cycle also corresponds
to the interval over which half of the star's surface is visible.  Of
the CQVs in Figures~\ref{fig:cqvs} and~\ref{fig:evoln}, a fraction that
seems greater than random might exhibit a preference for showing dips
or peaks that correspond to the half-cycle interval.

@@

In this work, we searched 120-second cadence TESS data collected
between July~2018 and Sep~2022 for complex quasiperiodic variables
(CQVs).  Our search sample included \nstarssearched\ K and M dwarfs
within 150\,pc, yielding a selection function with $\gtrsim 80\%$ completeness within 30\,pc,
and $\lesssim 10\%$ completeness at distances exceeding 100\,pc.

@@

The distinction between gas or dust is important because it will
clarify whether the CQV phenomenon is intrinsic, so that the material
comes from the star, or extrinsic, so that it is sourced through some
generic evolutionary phase of circumstellar disks.  This knowledge would in
turn propagate to our understanding of whether observing the
phenomenon is primarily teaching us about dust production and
processing in circumstellar disks, or whether it is teaching us about
the ability of cold gas to remain stable in hot stellar coronae for
long durations.
Observationally, acquisition of medium- or high-resolution time-series
spectroscopy holds the best chance at resolving the gas vs.~dust
question.  Given the observed evolutionary timescales, such data
should be acquired simultaneously with photometric time-series
observations to help clarify its interpretation.

@@

\begin{figure*}[!t]
	\begin{center}
    \includegraphics[width=0.45\textwidth]{ANNOTATED_resid_TIC_402980664_P18.5611_2min_phase_timegroups.png}
    	\end{center}
    \vspace{-0.4cm}
		\caption{
	      {\bf Alternative view of the evolution of LP 12-502}
	      (Figure~\ref{fig:lp}), arranged to emphasize changes in transit
	      times.  There are 200 binned black points per $P$=18.5611\,hr cycle; a two-harmonic
	      sinusoid has been subtracted from the PDC\_SAP light curve.
	      Vertical gray lines are underplotted to help guide the eye to instances
	      in which preferred dip phases synchronize over long baselines.
	      The orange and green lines guide the eye to where dips
	      change the positions of their local minima; orange
        lines have periods slightly less than $P$, the green line has
        a period greater than $P$.
		}
		\label{fig:lp2}
\end{figure*}

@@

\section{TIC~300651846}
\label{app:tic3006}
